### 2014-12-12 Review Response.md
Convert Review Response to Markdown

Response to reviewer comments

Reviewer #2
===========

Comment #1

Typo corrected as suggested

Comment #2

Typo corrected as suggested

Comment #3

A table has been added with benchmarks on memory use.

Comment #4

Typo corrected as suggested

Comment #5

Figure caption modified to explain notation of axes

Comment #6

Figure captions modified to explain labelling of lambda extrema and figures have been regenerated to improve the appearance of initial conditions

Comment #7

Typo corrected and sentence modified to improve grammar

Comment #8

Code listings have been revised and checked for poor breaks across lines and pages

Reviewer #3
===========

Comment #1

LCS Tool functions overview table added at beginning of numerical methods section.

Comment #2

Paragraph added at the beginning of Examples section to explain LCS Tool functions balance between performance and ease of use. Give approximate run-times for demo scripts.

Comment #3

The number of LCS detection software packages in the list of other LCS packages has been reduced and more details have been given. The “all-purpose” description of LCS Tool has been removed and the paragraph about LCS Tool's capabilities reworded.

Comment #4

A link to our article has been added to the LCS Tool README file.

Comment #5

The Highlights have been revised. The core contribution of the article is to present LCS Tool, a software tool to enable easy geodesic LCS computation. We believe the highlights explain this.

Reviewer #4
===========

The LCS Tool source code has been uploaded to RunMyCode, as suggested by the editor.

Reviewer #5
===========

Comment #1 and #2

The abstract and introduction have been revised based on the review comments.

Comment #3 and #4

Code listings have been revised and shortened.

Comment #6

LCS Tool functions overview table added at beginning of numerical methods section.